---
title: 'Categorical variables pre-lab: tidy'
author: "Professor McNamara"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

## Setup and packages

We use three packages in this course: `Lock5Data`, `tidyverse` and `infer`. To load a package, you use the `library()` function, wrapped around the name of a package. I've put the code to load one package into the chunk below. Add the other two you need. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(Lock5Data)
# put in the other two packages you need here
```

## Loading in data

Let's load the example dataset I have for this course, `GSS_clean.csv`. It should be inside the `data` folder in your RStudio Cloud. We'll use the `read_csv()` function to read in the data. Here is a chunk where we will put the data loading code. 

```{r data-load}
# this is where data-loading code goes
```

## Categorical variables

Categorical variables are those that describe a set of categories (e.g., "first year", "sophomore", "junior", "senior"), rather than a continuous numeric range. 

What variables in this dataset are categorical? How can you tell?

## Visualization for categorical variables

There are two main ways to visualize one categorical variable: **bar charts**, and pie charts. We're going to focus on bar charts here, because they are easier to create in R. 

When we generate graphics in the tidy R syntax, we use a main function called `ggplot()` (short for Grammar of Graphics plot, a reference to a famous book about data visualization). That function can take many "arguments", which are the things you put inside the parentheses. For example, if you were thinking about the algebra function $f(x)$, we would say the function is $f$, and the argument is $x$. In the case of `ggplot()`, the most important argument is the name of the dataset. Try running `ggplot(GSS)` and see what the result is. 

```{r}
# write code here
```

If we want an interesting plot, we need to use the `+` operator to add on a "geometric object" (or `geom_` for short) that tells R what kind of plot we want. In this case, we want a `geom_bar()` because we are making a bar chart. `geom_bar()` **also** takes arguments, and they have to be specified in a particular way. We use the `aes()` function (short for "aesthetics") and tell R how to map between a variable in our dataset and a variable in a plot. Here's an example:

```{r}
ggplot(GSS) + geom_bar(aes(x = marital_status))
```

It's important to get the syntax exactly right, or R will throw an error. (R is picky about capitalization, but not about spaces.)

Try making your own bar chart of the `labor_force_status`. 

```{r}
# write code here
```

### Two categorical variables
If we want, we can make a barchart showing **two** categorical variables. To do this, we add another argument to our `aes()` function. We will tell R to fill in the `marital_status` bars with color, according to the `born_in_us` variable. This makes a segmented or stacked barchart. 

```{r}
ggplot(GSS) + geom_bar(aes(x = marital_status, fill = born_in_us))
```

We can also make this barchart a side-by-side barchart by providing the `position` argument.

```{r}
ggplot(GSS) + geom_bar(aes(x = marital_status, fill = born_in_us), position = "dodge")
```

If we forgot what the argument was to make a side-by-side barchart, we could use R's built-in help functionality to assist us. 

```{r}
help(geom_bar)
```

Try making a barchart that shows both the `govt_or_private_employee` and `self_emp_or_works_for_somebody` variables. 

```{r}
# write code here
```

## Summary statistics for categorical variables

The other part of Exploratory Data Analysis (EDA) is making summary statistics. For a categorical variable, this means making frequency tables and relative frequency tables. Let's explore how to do that. 

For this piece, we will be using an operator called "the pipe", which is written as `%>%` (if you know other programming languages, you may think of `|` as the pipe, but that gets used to mean "or" in R). I think of the pipe as an arrow that points in the direction my data is flowing through a data pipeline. 

We will use the pipe to string together three main R functions:

- `summarize()`, which summarizes data (going from many rows down to just one or several)
- `mutate()`, which makes new variables in our dataset (sticking on one more column)
- `group_by()`, which works especially well with `summarize()` to group data for operations

These functions can be mixed and matched in many orders to create a multitude of data science products. 

Let's start with a short data pipeline.

```{r}
GSS %>%
  summarize(n = n())
```

What does this code tell us?

We can add another piece to the pipeline to make it more interesting.

```{r}
GSS %>%
  # another step
  summarize(n = n())
```

What does this code give us?

Finally, we might want to add one more piece to the pipeline to create a relative frequency table,

```{r}
GSS %>%
  # another step
  summarize(n = n()) %>%
  # one final step
```

### Two categorical variables

We can extend this to two categorical variables by adding a second variable into our `group_by()` code. Let's make a relative frequency table showing the relationship between `marital_status` and `general_happiness`. 

```{r}
GSS %>%
  group_by(marital_status) %>% # change one thing on this line
  summarize(n = n()) %>%
  mutate(prop = n/sum(n))
```

What happens if you switch the order of the variables in the `group_by()`?
