---
title: 'Inference for regression pre-lab: formula'
author: "Professor McNamara"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

## Setup and packages

We use three packages in this course: `Lock5Data`, `mosaic` and `ggformula`. To load a package, you use the `library()` function, wrapped around the name of a package. I've put the code to load one package into the chunk below. Add the other two you need. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(Lock5Data)
# put in the other two packages you need here
```


## Loading in data

As usual, we'll load the example data, `GSS_clean.csv`. It should be inside the `data` folder in your RStudio Cloud. We'll use the `read.csv()` function to read in the data. 

```{r data-load}
GSS <- read.csv("data/GSS_clean.csv")
```

## Regression

We've talked about regression before, in lab 3. However, that was just regression as a descriptive technique. Now we'll be going further to do inference about regression, using a distributional approximation. 

Let's consider the same relationship as we did before, predicting `highest_year_of_school_completed` (our response) using `highest_year_school_completed_spouse` (our explanatory variable). First, we should do a visualization of this relationship. 

```{r}

```

Now, let's compute the model and find the regression line

```{r}

```

We can use LaTeX to write out the equation of the line,

$$
\hat{y} = 5.85 + 0.59\cdot\text{highest year school completed spouse}
$$

Let's review how to interpret the slope cofficient:



## Inference for regression

R is great because it gives you all the inferential information you want in the `summary()` output,

```{r}

```

What is our test statistic? Our standard error?

What is our generic conclusion of the test? How can we interpret this in context?

What is the $R^2$ value? 

```{r}
rsquared(m1)
```

How do we interpret that in context?


## Confidence and prediction intervals

Like we did last time, we could make a prediction for a person whose spouse has completed 12 years of school (high school),

```{r}
5.84740 + 0.59403 * 12
```

but, we might want to create an interval estimate rather than just a point estimate. To do that, we will create either a confidence interval (for a mean Y) or a prediction interval (for a particular Y). Both of them can be created with the `predict()` function. 

```{r}
predict(m1, newdata = data.frame(highest_year_school_completed_spouse = 12), interval = "confidence")
predict(m1, newdata = data.frame(highest_year_school_completed_spouse = 12), interval = "prediction")
```

Let's interpret those two intervals



