---
title: 'Correlation and linear regression pre-lab: formula'
author: "Professor McNamara"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

## Setup and packages

We use three packages in this course: `Lock5Data`, `mosaic` and `ggformula`. To load a package, you use the `library()` function, wrapped around the name of a package. I've put the code to load one package into the chunk below. Add the other two you need. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(Lock5Data)
library(mosaic)
library(ggformula)
```

## Loading in data

Let's load the example dataset I have for this course, `GSS_clean.csv`. It should be inside the `data` folder in your RStudio Cloud. We'll use the `read.csv()` function to read in the data. 

```{r data-load}
GSS <- read.csv("data/GSS_clean.csv")
```

## Data visualization for two quantitative variables

Our main data visualization for two quantitative variables is a scatterplot. 

To make our visualization, we will use a function that begins with `gf_`. This time, the thing after the underscore is `point`, as in `gf_point()`. This function, like all functions in formula syntax, wants a formula as its first argument. 

Since we looking at a relationship between two variables, we use the $y~x$ syntax, which is the "formula." The `gf_` functions want another argument as well, which is the dataset where the variable is located. If you don't provide this, the function won't know where to look for the variables. Let's see an example,

```{r}
gf_point(highest_year_of_school_completed ~ highest_year_school_completed_spouse, data = GSS)
```

As with our plots in the last lab, we get a warning message about missing values being removed. 

Does this plot show a positive association, negative association, or no association? If it shows an association, how strong would you say it is?

## Summary statistics for two quantitative variables

One way to quantify the strength and direction of the relationship is with correlation. To find this, we use the `cor()` function, which takes standard formula syntax. 

```{r}
cor(highest_year_of_school_completed ~ highest_year_school_completed_spouse, data = GSS)
```

Much like with our other summary statistics, this gives an `NA` value, because there are `NA` values in one or more of the variables in the dataset. However, `cor()` doesn't use the standard `na.rm` argument, so we instead need to tell it what to `use`. In this case, we'd like to use "complete observations" (those where there is a value for both `highest_year_of_school_completed` and `highest_year_school_completed_spouse`), so we add `use = "complete.obs"` to our function arguments. 

```{r}
cor(highest_year_of_school_completed ~ highest_year_school_completed_spouse, data = GSS, use = "complete.obs")
```

## Linear models

Another way to study the relationship between two quantitative variables is to fit a linear model. R has a function that does least squares regression, called `lm()` for linear model. This also takes formula syntax. You can just run the function in your Console or RMarkdown document and get a little information about your model,

```{r}
lm(highest_year_of_school_completed ~ highest_year_school_completed_spouse, data = GSS)
```

but, it's better practice to assign your model object to a name, so you can refer back to it later, and get more information about it. Recall that the assignment operator in R is `<-`. I usually call my models `m1`, `m2`, etc., but you could think of a better name to use (like `schoolmodel` or similar) to remind you what it's about. 

```{r}
m1 <- lm(highest_year_of_school_completed ~ highest_year_school_completed_spouse, data = GSS)
```

When we run this code, nothing prints out. But, a new object should appear in our RStudio Environment pane. Now, we can run R functions on that model object. The most useful one is `summary()`,

```{r}
summary(m1)
```

The same model coefficients are shown here, but there is also a lot more information about things like the $R^2$ value. 

### Interpreting coefficients
Let's interpret our model coefficients. Here are the generic sentences I prefer:

**Intercept** If the value of [explanatory variable] was zero, our model would predict [response variable] to be [intercept value].

**Slope** For a one-[unit] increase in [explanatory variable], our model would predict a [slope value]-[unit] [increase/decrease] in [response variable]. 

Let's apply that to this model:




## Predicting values
Models are also useful for prediction. There are more programmatic ways to do prediction in R, but for now I recommend you just use R as a big calculator. Let's see what the model would predict for a person whose spouse had completed 12 years of school (high school):

```{r}
5.84740 + 0.59403 * 12
```

Does that make sense?

## Residuals
We are also interested in whether our model overpredicts or underpredicts certain points. We can compute a residual, which is $observed - expected$ or $y_i-\hat{y}_i$. Look at row 21 in the dataset. It represents a 34-year-old person with one child whose spouse completed 12 years of school. What is the observed value of `highest_year_of_school_completed` for this person? What is their residual?

```{r}

```




