---
title: 'Categorical variables pre-lab: formula'
author: "Professor McNamara"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

## Setup and packages

We use three packages in this course: `Lock5Data`, `mosaic` and `ggformula`. To load a package, you use the `library()` function, wrapped around the name of a package. I've put the code to load one package into the chunk below. Add the other two you need. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(Lock5Data)
# put in the other two packages you need here
```

## Loading in data

Let's load the example dataset I have for this course, `GSS_clean.csv`. It should be inside the `data` folder in your RStudio Cloud. We'll use the `read.csv()` function to read in the data. Here is a chunk where we will put the data loading code. 

```{r data-load}
# this is where data-loading code goes
```

## Categorical variables

Categorical variables are those that describe a set of categories (e.g., "first year", "sophomore", "junior", "senior"), rather than a continuous numeric range. 

What variables in this dataset are categorical? How can you tell?

## Visualization for categorical variables

There are two main ways to visualize one categorical variable: **bar charts**, and pie charts. We're going to focus on bar charts here, because they are easier to create in R. 

When we generate graphics in the formula R syntax, we use functions that begin with `gf_`. This stands for graphics formula (I think). After the underscore, there is the name of the plot we want to make. Because we want to make a barchart, we will use the `gf_bar()` function. This function, like all functions in formula syntax, wants a "formula" as it's first "argument." Let's talk about both those terms. 

First, an "argument" to a function is something that you put inside the parentheses. For example, if you were thinking about the algebra function $f(x)$, we would say the function is $f$, and the argument is $x$. In this case, the function is `gf_bar`, and we are going to give it a formula as an argument. 

Generally, formulas are of the format $y~x$. The y is the response variable, then there is a "tilde" (it's in the upper left of your keyboard, above the Tab key), then the explanatory variable(s). If you only have one variable, you don't give a $y$ and instead just give it $~x$. Let's see it in action.

```{r}
gf_bar(~marital_status, data = GSS)
```

Since we are only trying to visualize one variable, we use the tilde and then the name of the variable. That is the "formula." The `gf_` functions want another argument as well, which is the dataset where the variable is located. If you don't provide this, the function won't know where to look for `marital_status`. 

It's important to get the syntax exactly right, or R will throw an error. (R is picky about capitalization, but not about spaces.)

Try making your own bar chart of the `labor_force_status`. 

```{r}
# write code here
```

### Two categorical variables
If we want, we can make a barchart showing **two** categorical variables. To do this, we add another argument to our `gf_bar` function. We will tell R to fill in the `marital_status` bars with color, according to the `born_in_us` variable. This makes a segmented or stacked barchart. 

```{r}
gf_bar(~marital_status, fill = ~born_in_us, data = GSS)
```

Notice that again, when we refer to the variable `born_in_us`, we have to use the tilde to make it a formula. 

We can also make this barchart a side-by-side barchart by providing the `position` argument.

```{r}
gf_bar(~marital_status, fill = ~born_in_us, data = GSS, position = "dodge")
```

Try making a barchart that shows both the `govt_or_private_employee` and `self_emp_or_works_for_somebody` variables. 

```{r}
# write code here
```

## Summary statistics for categorical variables

The other part of Exploratory Data Analysis (EDA) is making summary statistics. For a categorical variable, this means making frequency tables and relative frequency tables. Let's explore how to do that. 

Again, we will be using formula notation. In this case, we will use the `tally()` function. Its arguments are almost the same as `gf_bar()`. 

```{r}
tally(~marital_status, data = GSS)
```

What does this code tell us?

We can add another argument to make it more interesting.

```{r}
tally(~marital_status, data = GSS, format = "proportion")
```

What does this code give us?


### Two categorical variables

We can extend this to two categorical variables by adding a second variable into our code. Let's make a relative frequency table showing the relationship between `marital_status` and `general_happiness`. 

```{r}
tally(marital_status ~ general_happiness, data = GSS, format = "proportion")
```

What happens if you switch the order of the variables in the formula?
